---
title: "Gradients FA Summary"
author: "CTELab"
format: 
  revealjs:
    theme: night
editor: visual
incremental: true
---

```{r}

library(tidyverse) # data cleanup
library(viridis) # color palette
library(ggfortify) # PCA visualizations
library(ggforce) # added ggplot functions
library(vegan) # multivariate analyses

# load in all data
all_species <- read_csv("Data/Biomarkers/FattyAcids/gradients2019_corespecies_FA_QAQC.csv", 
    col_types = cols(`19:0` = col_skip()))
# subset inverts
all_inverts <- all_species %>%
          filter(type == "invert")
# subset algae
all_algae <- all_species %>%
          filter(type == "algae")
# Create simplified long dataset for analysis of algae
long_algae <- all_algae %>%
  select(FAsampleName, `NIC-Klein-Midpoint-Annual`, genusSpecies, `8:0`:`22:4w3`) %>%
  pivot_longer(cols = `8:0`:`22:4w3`, names_to = 'FA', values_to = 'proportion')
 # pivot algae data wide for PCA/nMDS
 grad_conc_wide <- long_algae %>%
   select(FA, genusSpecies, proportion, `NIC-Klein-Midpoint-Annual`, FAsampleName) %>%
   pivot_wider(names_from = FA, values_from = proportion, values_fill = 0)
 # Create simplified long dataset for analysis of inverts
long_inverts <- all_inverts %>%
  select(FAsampleName, genusSpecies, `NIC-Klein-Midpoint-Annual`, `8:0`:`22:4w3`) %>%
  pivot_longer(cols = `8:0`:`22:4w3`, names_to = 'FA', values_to = 'proportion')
 # pivot invert data wide for PCA/nMDS
 grad_invert_wide <- long_inverts %>%
   select(FA, genusSpecies, proportion, FAsampleName, `NIC-Klein-Midpoint-Annual`) %>%
   pivot_wider(names_from = FA, values_from = proportion, values_fill = 0)

```

## Sample Processing

-   Invertebrate Samples

    -   559 of 559 complete

    -   15 sites

    -   11 species

    -   

        ```{r}

        list(unique(all_inverts$genusSpecies))

        ```

## Sample Processing

-   Algal Samples

    -   106 of 155 complete

        -   remaining chromatograms in progress

    -   13 sites

    -   7 species

    -   

        ```{r}

        list(unique(all_algae$genusSpecies))

        ```

## Preliminary Results - Algae

## FA Proportions - Algae

```{r}

# dotplot of final concentration by species by FA
ggplot(filter(long_algae, FA %in% c("14:0", "16:0", "16:1w7c", "18:0", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "20:4w6", "20:5w3")), 
       aes(x = FA, y = proportion, colour = genusSpecies)) +
  geom_point(size = 4, position = position_jitter(width = .1)) +
  geom_boxplot(data = long_algae %>% 
                 filter(FA %in% c("14:0", "16:0", "16:1w7c", "18:0", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "20:4w6", "20:5w3")), 
               aes(group = FA, y = proportion), color = "black", alpha = 0, show.legend = FALSE) +
  theme_classic() +
  scale_colour_viridis(discrete = TRUE, end = 0.9) +
  labs(x = "Fatty Acid", y = "Proportion of Total Fatty Acids") +
  scale_x_discrete(labels = c("14:0" = "MYR (14:0)",
                              "16:0" = "PAL (16:0)", 
                            "16:1w7c" = "PALO (16:1w7c)", 
                            "18:0" = "STE (18:0)",
                            "18:1w7c" = "VAC (18:1w7c)",
                            "18:3w3" = "LIN (18:3w3)",
                            "18:4w3c" = "SDA (18:4w3c)",
                            "18:1w9c" = "OLE (18:1w9c)", 
                            "20:4w6" = "ARA (20:4w6)", 
                            "20:5w3" = "EPA (20:5w3)")) +
  coord_flip()

```

## Mean FA Composition - Algae

```{r}

# stacked barplot of percent contribution of each FA to total FA (dominant)
ggplot(filter(long_algae, FA %in% c("14:0", "16:0", "16:1w7c", "18:0", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "20:4w6", "20:5w3")) %>%
         group_by(genusSpecies, FA) %>%
         summarise(proportion = mean(proportion)) %>%
         ungroup(),
         aes(x = proportion, y = genusSpecies, fill = factor(FA, levels = c("14:0", "18:0", "16:1w7c", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "16:0", "20:5w3", "20:4w6")))) +
  geom_col(position = "stack", color = "black") +
  scale_fill_viridis(discrete = TRUE, option = "H") +
  theme_classic() +
  labs(fill = "Fatty Acid") +
  xlab("Mean Proportion of Total Fatty Acids") +
  ylab("Species")

```

## nMDS - Algae

```{r, error = FALSE, warning = FALSE, message = FALSE}

# select EPA, ARA, SDA, PAL, OLE, LIN, VAC, and dominant sats (16, 18) 
sub_wide <- grad_conc_wide %>% # fix 16:0 error in inverts
  select(genusSpecies, FAsampleName, "14:0", "16:0", "16:1w7c", "18:0", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "20:4w6", "20:5w3")
sub_wide_trans <- (sub_wide[,3:12])
sub_wide_trans <- bind_cols(sub_wide[1:2], sub_wide_trans)

batch_1_2_MDS <- metaMDS(sub_wide_trans[3:12], autotransform = TRUE, distance = "clark", trace = FALSE)
batch_1_2_MDS_points <- batch_1_2_MDS$points
batch_1_2_MDS_points <- data.frame(batch_1_2_MDS_points)
plot_data_batch_1_2 <- data.frame(batch_1_2_MDS_points, sub_wide[,1])

plot_data_batch_1_2 %>%
ggplot(aes(MDS1, MDS2, color = genusSpecies)) +
  geom_point(size = 4) + 
  annotate(geom = "text", x = 1, y = 1, label = "2D stress = 0.17", color = 'grey') +
  scale_color_viridis(discrete = TRUE, end = 0.9, option = 'D') + 
   geom_mark_hull(expand = unit(1, "mm"), radius = unit(1, "mm"), concavity = 5)  + 
  theme_void() 

```

## nMDS - Algae

```{r}
# PERMANOVA 

 # remove zero columns
 grad_conc_adonis <- grad_conc_wide %>%
   select(where(~ any(. != 0)))
 
 # algal FA for adonis
 FA_only <- grad_conc_adonis %>%
   select(`8:0`:`18:3w1`)

adonis(abs(FA_only) ~ genusSpecies * as.character(`NIC-Klein-Midpoint-Annual`), data = grad_conc_adonis, method = 'bray', perm = 9999)

```

## PCA - Algae

```{r, error = FALSE, warning = FALSE, message = FALSE}

 # remove zero columns
 grad_conc_PCA <- grad_conc_wide %>%
   select(where(~ any(. != 0)))
 # run PCA
PCA_results <-  prcomp(grad_conc_PCA[,c(4:66)], scale = TRUE)
PCA_results$rotation <- -1*PCA_results$rotation
#reverse the signs of the scores
PCA_results$x <- -1*PCA_results$x



autoplot(PCA_results, data = grad_conc_PCA, colour = 'genusSpecies', 
         size = 4,
         loadings = TRUE, 
         loadings.colour = 'grey', 
         loadings.label = TRUE, 
         loadings.label.size = 4) +
  scale_color_viridis(discrete = TRUE) +
  theme_minimal()


```

## Species Gradients - Algae

```{r, error = FALSE, warning = FALSE, message = FALSE}

 # remove zero columns
 grad_conc_PCA <- grad_conc_wide %>%
   select(where(~ any(. != 0)))
# filter to algae and FA of interest (top 20 most common b/w species)
grad_conc_alga <- grad_conc_PCA %>%
  filter(genusSpecies %in% c("Desmarestia menziesii", "Phyllophora antarctica", "Himantothallus grandifolius")) %>%
  select(genusSpecies, `NIC-Klein-Midpoint-Annual`,
         `20:4w6`,
         `20:5w3`,
         `16:0`,
         `18:1w9c`,
         `20:5w3`,
         `20:4w6`,
         `18:3w3`,
         `16:1w7c`,
         `18:1w7c`,
         `18:3w3`,
         `18:2w6c`,
         `14:0`,
         `18:4w1`,
         `18:4w3c`,
         `18:0`,
         `22:5w3`,
         `20:0`,
         `20:4w3`,
         `14:0`,
         `22:4w6`)
grad_conc_alga <- grad_conc_alga %>%
  mutate(gradient = as.character(`NIC-Klein-Midpoint-Annual`))

# run PCA
PCA_results <-  prcomp(grad_conc_alga[,c(3:18)], scale = TRUE)
PCA_results$rotation <- -1*PCA_results$rotation
#reverse the signs of the scores
PCA_results$x <- -1*PCA_results$x

autoplot(PCA_results, data = grad_conc_alga, colour = 'gradient', shape = 'genusSpecies',
         size = 4,
         loadings = TRUE, 
         loadings.colour = 'grey', 
         loadings.label = TRUE, 
         loadings.label.size = 4) +
 scale_color_viridis(discrete = TRUE, option = "F", begin = 0.2, end = 0.9) +
  theme_minimal()


```

## Specific FA Gradient - Algae

```{r, fig.width=5.2, fig.height=7, fig.align='center'}

algal_grad <- grad_conc_wide %>%
  select(genusSpecies, `NIC-Klein-Midpoint-Annual`, '14:0', '16:0', '16:1w7c', '18:0', '18:4w3c', '20:4w6', '20:5w3') %>%
  filter(genusSpecies %in% c("Desmarestia menziesii", "Phyllophora antarctica", "Himantothallus grandifolius"))
algal_grad_long <- algal_grad %>%
  pivot_longer(cols = `14:0`:`20:5w3`, names_to = 'FA', values_to = 'proportion')

  
algal_grad_long %>%
  ggplot(aes(`NIC-Klein-Midpoint-Annual`, proportion)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_grid(FA ~ genusSpecies, scales = 'free') +
  theme_bw()
  
  

```

## Preliminary Results - Inverts (non-ANTSEA)

## FA Proportions - Inverts

```{r}

# dotplot of final concentration by species by FA
ggplot(filter(long_inverts, FA %in% c("14:0", "16:0", "16:1w7c", "18:0", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "20:4w6", "20:5w3")), 
       aes(x = FA, y = proportion, colour = genusSpecies)) +
  geom_point(size = 1, position = position_jitter(width = .1)) +
  geom_boxplot(data = long_inverts %>% 
                 filter(FA %in% c("14:0", "16:0", "16:1w7c", "18:0", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "20:4w6", "20:5w3")), 
               aes(group = FA, y = proportion), color = "black", alpha = 0, show.legend = FALSE) +
  theme_classic() +
  scale_colour_viridis(discrete = TRUE, end = 0.9, option = "C") +
  labs(x = "Fatty Acid", y = "Proportion of Total Fatty Acids") +
  scale_x_discrete(labels = c("14:0" = "MYR (14:0)",
                              "16:0" = "PAL (16:0)", 
                            "16:1w7c" = "PALO (16:1w7c)", 
                            "18:0" = "STE (18:0)",
                            "18:1w7c" = "VAC (18:1w7c)",
                            "18:3w3" = "LIN (18:3w3)",
                            "18:4w3c" = "SDA (18:4w3c)",
                            "18:1w9c" = "OLE (18:1w9c)", 
                            "20:4w6" = "ARA (20:4w6)", 
                            "20:5w3" = "EPA (20:5w3)")) +
  coord_flip()

```

## Mean FA Composition - Inverts

```{r}

# stacked barplot of percent contribution of each FA to total FA (dominant)
ggplot(filter(long_inverts, FA %in% c("14:0", "16:0", "16:1w7c", "18:0", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "20:4w6", "20:5w3")) %>%
         group_by(genusSpecies, FA) %>%
         summarise(proportion = mean(proportion)) %>%
         ungroup(),
         aes(x = proportion, y = genusSpecies, fill = factor(FA, levels = c("14:0", "18:0", "16:1w7c", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "16:0", "20:5w3", "20:4w6")))) +
  geom_col(position = "stack", color = "black") +
  scale_fill_viridis(discrete = TRUE, option = "H") +
  theme_classic() +
  labs(fill = "Fatty Acid") +
  xlab("Mean Proportion of Total Fatty Acids") +
  ylab("Species")

```

## nMDS - Inverts

```{r, error = FALSE, warning = FALSE, message = FALSE}

# select EPA, ARA, SDA, PAL, OLE, LIN, VAC, and dominant sats (16, 18) 
sub_wide <- grad_invert_wide %>% # fix 16:0 error in inverts
  select(genusSpecies, FAsampleName, "14:0", "16:0", "16:1w7c", "18:0", "18:1w7c", "18:3w3", "18:4w3c", "18:1w9c", "20:4w6", "20:5w3")
sub_wide_trans <- (sub_wide[,3:12])
sub_wide_trans <- bind_cols(sub_wide[1:2], sub_wide_trans)

batch_1_2_MDS <- metaMDS(sub_wide_trans[3:12], autotransform = TRUE, distance = "clark", trace = FALSE)
batch_1_2_MDS_points <- batch_1_2_MDS$points
batch_1_2_MDS_points <- data.frame(batch_1_2_MDS_points)
plot_data_batch_1_2 <- data.frame(batch_1_2_MDS_points, sub_wide[,1])


plot_data_batch_1_2 %>%
ggplot(aes(MDS1, MDS2, color = genusSpecies)) +
  geom_point(size = 2) + 
  annotate(geom = "text", x = 0.75, y = 1, label = "2D stress > 0.20", color = 'grey') +
  scale_color_viridis(discrete = TRUE, end = 0.9, option = 'C') +
   geom_mark_hull(expand = unit(1, "mm"), radius = unit(1, "mm"), concavity = 5)  + 
  theme_void() 



```

## nMDS - Inverts

```{r}

# PERMANOVA 

 # remove zero columns
 grad_conc_adonis <- grad_invert_wide %>%
   select(where(~ any(. != 0)))
 
 # algal FA for adonis
 FA_only <- grad_conc_adonis %>%
   select(`13:0`:`22:4w3`)

adonis(abs(FA_only) ~ genusSpecies * as.character(`NIC-Klein-Midpoint-Annual`), data = grad_conc_adonis, method = 'bray', perm = 9999)

```

## PCA - Inverts

```{r, error = FALSE, warning = FALSE, message = FALSE}

 # remove zero columns
 grad_conc_PCA <- grad_invert_wide %>%
   select(where(~ any(. != 0)))
 # run PCA
PCA_results <-  prcomp(grad_conc_PCA[,c(4:65)], scale = TRUE)
PCA_results$rotation <- -1*PCA_results$rotation
#reverse the signs of the scores
PCA_results$x <- -1*PCA_results$x



autoplot(PCA_results, data = grad_conc_PCA, colour = 'genusSpecies', 
         size = 4,
         loadings = TRUE, 
         loadings.colour = 'grey', 
         loadings.label = TRUE, 
         loadings.label.size = 4) +
  scale_color_viridis(discrete = TRUE, option = "C") +
  theme_minimal()


```

## Specific FA Gradient - Inverts

```{r, fig.width=5.2, fig.height=7, fig.align='center'}

invert_grad <- grad_invert_wide %>%
  select(genusSpecies, `NIC-Klein-Midpoint-Annual`, '14:0', '16:0', '16:1w7c', '18:0', '18:3w6', '20:0', '20:4w6', '22:0') %>%
  filter(genusSpecies %in% c("Prostebbingia gracilis", "Dendrilla membranosa", "Gondogeneia antarctica"))
invert_grad_long <- invert_grad %>%
  pivot_longer(cols = `14:0`:`22:0`, names_to = 'FA', values_to = 'proportion')

  
invert_grad_long %>%
  ggplot(aes(`NIC-Klein-Midpoint-Annual`, proportion)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_grid(FA ~ genusSpecies, scales = 'free') +
  theme_bw()


```

## Invert Species Gradients - Amphipods

```{r, error = FALSE, warning = FALSE, message = FALSE}

 # remove zero columns
 grad_conc_PCA <- grad_invert_wide %>%
   select(where(~ any(. != 0)))
# filter to amphipods and FA of interest (top 20 most common b/w species)
grad_conc_amph <- grad_conc_PCA %>%
  filter(genusSpecies %in% c("Prostebbingia gracilis", "Gondogeneia antarctica")) %>%
  select(genusSpecies, `NIC-Klein-Midpoint-Annual`,
         `16:0`,
         `20:5w3`,
         `22:6w3`,
         `18:1w9c`,
         `16:4w1`,
         `18:4w3`,
         `14:0`,
         `16:1w7c`,
         `18:1w7c`,
         `18:3w3`,
         `16:3w4`,
         `18:0`,
         `16:2w6`,
         `18:2w6c`,
         `20:4w6`,
         `16:4w3`,
         `20:4w3`,
         `20:1w9`,
         `20:3w3`,
         `22:5w3`)
grad_conc_amph <- grad_conc_amph %>%
  mutate(gradient = as.character(`NIC-Klein-Midpoint-Annual`))

# run PCA
PCA_results <-  prcomp(grad_conc_amph[,c(3:22)], scale = TRUE)
PCA_results$rotation <- -1*PCA_results$rotation
#reverse the signs of the scores
PCA_results$x <- -1*PCA_results$x

autoplot(PCA_results, data = grad_conc_amph, colour = 'gradient', shape = 'genusSpecies',
         size = 4,
         loadings = TRUE, 
         loadings.colour = 'grey', 
         loadings.label = TRUE, 
         loadings.label.size = 4) +
 scale_color_viridis(discrete = TRUE, option = "F", begin = 0.2, end = 0.9) +
  theme_minimal()


```

## Invert Species Gradients - Dendrilla

```{r, error = FALSE, warning = FALSE, message = FALSE}

 # remove zero columns
 grad_conc_PCA <- grad_invert_wide %>%
   select(where(~ any(. != 0)))
# filter to Dendrilla and FA of interest (top 14 most common b/w species)
grad_conc_dend <- grad_conc_PCA %>%
  filter(genusSpecies %in% c("Dendrilla membranosa")) %>%
  select(genusSpecies, `NIC-Klein-Midpoint-Annual`,
         `22:6w3`,
         `16:0`,
         `22:5w3`,
         `20:5w3`,
         `18:0`,
         `14:0`,
         `22:0`,
         `20:4w3`,
         `18:4w3`,
         `i-15:0`,
         `18:1w7c`,
         `22:4w3`,
         `20:0`,
         `a-15:0`)
grad_conc_dend <- grad_conc_dend %>%
  mutate(gradient = as.character(`NIC-Klein-Midpoint-Annual`))


# run PCA
PCA_results <-  prcomp(grad_conc_dend[,c(3:16)], scale = TRUE)
PCA_results$rotation <- -1*PCA_results$rotation
#reverse the signs of the scores
PCA_results$x <- -1*PCA_results$x



autoplot(PCA_results, data = grad_conc_dend, colour = 'gradient', shape = 'genusSpecies',
         size = 4,
         loadings = TRUE, 
         loadings.colour = 'grey', 
         loadings.label = TRUE, 
         loadings.label.size = 4) +
  scale_color_viridis(discrete = TRUE, option = "F", begin = 0.2, end = 0.9) +
  theme_minimal()


```

## Next Steps

-   Finish algal chromatogram analysis (\~3 weeks)

-   Complete data pipeline (\~2 weeks)

    -   finish species/site updates

-   Continue exploratory analyses

    -   SAFA vs. MUFA/PUFA proportions

-   Choose strongest/most interesting detected patterns

-   PUBLISH!

## 
