---
title: "Biomarker Sample Summary"
author: "Ross Whippo"
date: "10/17/2019"
output: 
  tufte::tufte_handout: default
  tufte::tufte_html: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(tufte)
library(kableExtra)
```


## Overview of samples collected

```{r, echo = FALSE, message = FALSE}
samples <- read_csv("B-236_Fatty-Acid_Collections_QAQC.csv")
```

\newthought{Over the course} of the expedition, a total of `r length(samples$ProjID)` unique samples were collected including members of `r length(unique(samples$phylum, na.rm = TRUE))` phyla, `r length(unique(samples$class, na.rm = TRUE))` classes, `r length(unique(samples$Genus, na.rm = TRUE))` genera, and `r length(unique(samples$genusSpecies, na.rm = TRUE))` species\footnote{Currently, organisms that are not identified to the species level result in the discrepancy between number of genera and species.}.

$~$

**Phyla**
```{r, echo = FALSE}
phyla <- unique(samples$phylum)
phyla <- phyla %>%
  na.omit()
unique(phyla)
```

**Classes**
```{r, echo = FALSE}
class <- unique(samples$class)
class <- class %>%
  na.omit()
unique(class)
```

```{r, echo = FALSE, message = FALSE}
all_inverts <- samples %>%
  filter(kingdom == "Animalia" | is.na(kingdom))
all_algae <- samples %>%
  filter(kingdom %in% c('Plantae', 'Chromista'))
```
A total of `r length(all_inverts$ProjID)` invertebrate and `r length(all_algae$ProjID)` algal samples from 15 different sites were collected\footnote{These numbers are approximate due to issues identified above.}.

$~$

Number of specimens collected varied by specific site, latitude, and ice cover. 

```{r, echo = FALSE, fig.margin = TRUE, warning = FALSE}
samples$count <- 1
ggplot(samples, aes(x = SiteID, y = count)) +
         geom_col()
```

```{r, echo = FALSE, fig.margin = TRUE, warning = FALSE}
samples$count <- 1
ggplot(samples, aes(x = IceCoverCat, y = count)) +
         geom_col()
```

## Summary of invertebrates collected

\newthought{Number of invertebrates} collected varied by specific site, latitude, and ice cover parameters. The greatest number of samples were collected at: sites 05, 07:12, and 14:15 (>100 samples); ice cover categories 0.4, 0.6, and 0.8 (>300 samples); and at latitudes to be determined. 


Invertebrates present with at least two replicates at all sites include *Nacella concinna* and *Odontaster validus*. 

\newthought{The following table} lists all collected invertebrates with the number of replicates for the number of sites with that replicate size. 

```{r, echo = FALSE, fig.margin = TRUE, message = FALSE, warning = FALSE}
library(dplyr)
sitereps <- read_csv('core_species_siterep.csv')
siterepkable <- kable(sitereps[2:4], longtable = TRUE) 
siterepkable
```

## Summary of algae collected

```{r, echo = FALSE, messae = FALSE, warning = FALSE}
algaesamps <- samples %>%
  filter(phylum %in% c("Rhodophyta", "Chlorophyta", "Ochrophyta")) %>%
  group_by(genusSpecies)
algaltable <- algaesamps %>%
  select(phylum, genusSpecies) %>%
    arrange(desc(phylum)) %>%
  distinct(phylum, genusSpecies) %>%

  kable()
algaltable
  
```

You can also include things in the margin that are not numbered with the margin note command\marginnote{Like this.}.

```
this is inline code.
```
