---
title: "Biomarker Sample Summary"
author: "Ross Whippo"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
output: 
  tufte::tufte_handout: default
  tufte::tufte_html: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(tufte)
library(kableExtra)
```


## Overview of samples collected

```{r, echo = FALSE, message = FALSE}
samples <- read_csv("B-236_Fatty-Acid_Collections_QAQC.csv")
```

\newthought{Over the course} of the expedition, a total of `r length(samples$ProjID)` unique samples were collected including members of `r length(unique(samples$phylum, na.rm = TRUE))` phyla, `r length(unique(samples$class, na.rm = TRUE))` classes, `r length(unique(samples$Genus, na.rm = TRUE))` genera, and `r length(unique(samples$genusSpecies, na.rm = TRUE))` species\footnote{Currently, organisms that are not identified to the species level result in the discrepancy between number of genera and species.}.

$~$

The Phyla and Classes collected include:

$~$

**Phyla**
```{r, echo = FALSE}
phyla <- unique(samples$phylum)
phyla <- phyla %>%
  na.omit()
unique(phyla)
```


```{r, echo = FALSE, fig.margin = TRUE, warning = FALSE}
samples$count <- 1
samples$LAT_DD.1 <- as.factor(samples$LAT_DD.1)
ggplot(samples, aes(x = LAT_DD.1, y = count)) +
         geom_col() + 
  theme(axis.text.x = element_text(angle = 270))
          
```
```{marginfigure}
These figures show the number of samples collected (count) by latitude and ice cover.
```
```{r, echo = FALSE, fig.margin = TRUE, warning = FALSE}
samples$count <- 1
ggplot(samples, aes(x = IceCoverCat, y = count)) +
         geom_col()
```


**Classes**
```{r, echo = FALSE}
class <- unique(samples$class)
class <- class %>%
  na.omit()
unique(class)
```

```{r, echo = FALSE, message = FALSE}
all_inverts <- samples %>%
  filter(kingdom == "Animalia" | is.na(kingdom))
all_algae <- samples %>%
  filter(kingdom %in% c('Plantae', 'Chromista'))
```
A total of `r length(all_inverts$ProjID)` invertebrate and `r length(all_algae$ProjID)` algal samples from 15 different sites were collected\footnote{These numbers are approximate due to issues identified above.}. The most abundnant organisms in terms of replicates include:
```{r, echo = FALSE}
most_taxa <- samples %>%
  select(count, genusSpecies) %>%
  group_by(genusSpecies) %>%
  summarise(sum(count)) 
names(most_taxa)[names(most_taxa)=="sum(count)"] <- "count"
arrange(most_taxa, desc(count))
```


$~$

Number of specimens collected varied by specific site, latitude, and ice cover as seen in figures \ref{fig:allsite}, \ref{fig:alllat}, and \ref{fig:allice}. 

```{r, echo = FALSE, fig.fullwidth = TRUE, message = FALSE, warning = FALSE, fig.cap = "Number of samples collected per site by phylum.\\label{fig:allsite}"}
library(viridis)
ggplot(data = subset(samples, !is.na(phylum)), aes(x = SiteID, y = count)) +
  geom_col(aes(fill = phylum)) +
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "right", text = element_text(size = 8)) +
  guides(shape = guide_legend(override.aes = list(size = 1))) +
  theme(legend.title = element_text(size = 3), 
               legend.text = element_text(size = 4))
```


```{r, echo = FALSE, fig.fullwidth = TRUE, warning = FALSE, message = FALSE, fig.height = 20, fig.width = 15, fig.cap = "Heat map of number of reps per species collected by latitude.\\label{fig:alllat}"}
all_heat <- samples
# for latitude tables
all_heat_lat <- all_heat %>%
  group_by(LAT_DD.1, genusSpecies) %>%
  summarize(sum(count))
# rename column
names(all_heat_lat)[names(all_heat_lat)=="sum(count)"] <- "totalSpecies"
all_heat_lat$LAT_DD.1 <- as.character(all_heat_lat$LAT_DD.1)
ggplot(data = all_heat_lat, aes(x = LAT_DD.1, y = genusSpecies)) +
  geom_tile(aes(fill = totalSpecies)) +
  scale_fill_viridis(option = "B", begin = 0.3, end = 1) +
  theme_dark() +
  theme(panel.background = element_rect(fill = 'black')) +
  theme(text = element_text(size = 18)) +
  geom_text(show.legend = FALSE, colour = "white", size = 6, aes(label = totalSpecies)) + 
  theme(axis.text.x = element_text(angle = 270))

```

```{r, echo = FALSE, fig.fullwidth = TRUE, warning = FALSE, message = FALSE, fig.height = 20, fig.width = 15, fig.cap = "Heat map of number of reps per species collected by ice cover category.\\label{fig:allice}"}
all_heat <- samples
# for ice tables
all_heat_cover <- all_heat %>%
  group_by(IceCoverCat, genusSpecies) %>%
  summarize(sum(count))
# rename column
names(all_heat_cover)[names(all_heat_cover)=="sum(count)"] <- "totalSpecies"
ggplot(data = all_heat_cover, aes(x = IceCoverCat, y = genusSpecies)) +
  geom_tile(aes(fill = totalSpecies)) +
  scale_fill_viridis(option = "B", begin = 0.3, end = 1) +
  theme_dark() +
  theme(panel.background = element_rect(fill = 'black')) +
  theme(text = element_text(size = 18)) +
  geom_text(show.legend = FALSE, colour = "white", size = 6, aes(label = totalSpecies)) + 
  theme(axis.text.x = element_text(angle = 270))

```

\newthought{The following table} lists all collected organisms with the number of replicates for the number of sites with that replicate size. 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
sitereps <- read_csv('core_species_siterep.csv')
siterepkable <- kable(sitereps[2:4], longtable = TRUE) 
siterepkable
```

## Summary of invertebrates collected

\newthought{Number of invertebrates} collected varied by specific site, latitude, and ice cover parameters. The greatest number of samples were collected at: sites 05, 07:12, and 14:15 (>100 samples); ice cover categories 0.4, 0.6, and 0.8 (>300 samples); and at lower latitudes with the majority of sites below 65$^{\circ}$ each having >100 samples each. 

$~$

Invertebrates present with at least two replicates at all sites include *Nacella concinna* and *Odontaster validus*\footnote{\textit{Cnemidocarpa} sp. has at least two (2) replicates for all but one (1) site}. The most commonly sampled invertebrates include:

```{r, echo = FALSE}
most_inverts <- samples %>%
  filter(kingdom == "Animalia" | is.na(kingdom)) %>%
  select(count, genusSpecies) %>%
  group_by(genusSpecies) %>%
  summarise(sum(count)) 
names(most_inverts)[names(most_inverts)=="sum(count)"] <- "count"
arrange(most_inverts, desc(count))
```


Roughly 13 invertebrate species have a mimimum required number of reps (3) across six (6) sites. These numbers decrease as you increase the number of reps or number of sites desired. Invertebrates tended to be distributed more evenly across all gradients than algae (see below). Distribution of invertebrates by latitude and ice cover can be found in figures \ref{fig:invertlat} and \ref{fig:invertice} respectively. 

```{r, echo = FALSE, fig.fullwidth = TRUE, warning = FALSE, message = FALSE, fig.height = 20, fig.width = 15, fig.cap = "Heat map of number of reps per invertebrate species collected by latitude.\\label{fig:invertlat}"}
invert_heat <- samples %>%
  filter(kingdom == "Animalia")
invert_heat_lat <- invert_heat %>%
  group_by(LAT_DD.1, genusSpecies) %>%
  summarize(sum(count))
# rename column
names(invert_heat_lat)[names(invert_heat_lat)=="sum(count)"] <- "totalSpecies"
ggplot(data = invert_heat_lat, aes(x = LAT_DD.1, y = genusSpecies)) +
  geom_tile(aes(fill = totalSpecies)) +
  scale_fill_viridis(option = "B", begin = 0.3, end = 1) +
  theme_dark() +
  theme(panel.background = element_rect(fill = 'black')) +
  theme(text = element_text(size = 22)) +
  geom_text(show.legend = FALSE, colour = "white", size = 8, aes(label = totalSpecies)) + 
  theme(axis.text.x = element_text(angle = 270))

  ```

```{r, echo = FALSE, fig.fullwidth = TRUE, warning = FALSE, message = FALSE, fig.height = 20, fig.width = 15, fig.cap = "Heat map of number of reps per invertebrate species collected by ice cover category.\\label{fig:invertice}"}
invert_heat <- filter(samples, kingdom == "Animalia" | is.na(kingdom), genusSpecies != "Benthic diatoms")
# for ice cover heat maps 
invert_heat_cover <- invert_heat %>%
  group_by(IceCoverCat, genusSpecies) %>%
  summarize(sum(count))
# rename column
names(invert_heat_cover)[names(invert_heat_cover)=="sum(count)"] <- "totalSpecies"
ggplot(data = invert_heat_cover, aes(x = IceCoverCat, y = genusSpecies)) +
  geom_tile(aes(fill = totalSpecies)) +
  scale_fill_viridis(option = "B", begin = 0.3, end = 1) +
  theme_dark() +
  theme(panel.background = element_rect(fill = 'black')) +
  theme(text = element_text(size = 22)) +
  geom_text(show.legend = FALSE, colour = "white", size = 8, aes(label = totalSpecies)) + 
  theme(axis.text.x = element_text(angle = 270))

```

\pagebreak

## Summary of algae collected

\newthought{Algal collections} comprised roughly `r round(length(all_algae$ProjID)/length(samples$ProjID)*100)`% of all samples taken. 

```{r, echo = FALSE}
most_algae <- samples %>%
  filter(kingdom %in% c('Plantae', 'Chromista')) %>%
  select(count, genusSpecies) %>%
  group_by(genusSpecies) %>%
  summarise(sum(count)) 
names(most_algae)[names(most_algae)=="sum(count)"] <- "count"
arrange(most_algae, desc(count))
```



```{r, echo = FALSE, message = FALSE, warning = FALSE}
algaesamps <- samples %>%
  filter(phylum %in% c("Rhodophyta", "Chlorophyta", "Ochrophyta")) %>%
  group_by(genusSpecies)
algaltable <- algaesamps %>%
  select(phylum, genusSpecies) %>%
    arrange(desc(phylum)) %>%
  distinct(phylum, genusSpecies)
```

\newthought{A total of} `r length(algaltable$phylum)` algal species were collected, with the majority of algae representing the phylum Rhodophyta (~`r round(nrow(filter(algaltable, phylum == "Rhodophyta"))/length(algaltable$phylum)*100)`%), and only one representative of the Chorohyta, with the remainder from the Ochrophyta. Algal diversity was concentrated in the lower latitudes and lower ice cover zones as seen in figures \ref{fig:algallat} and \ref{fig:algalice} respectively. 

$~$

This table summarises the identity of all algal species collected and their respective phyla.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
kable_algae <-kable(algaltable, longtable = TRUE)
kable_algae
```

```{r, echo = FALSE, fig.fullwidth = TRUE, warning = FALSE, message = FALSE, fig.height = 20, fig.width = 15, fig.cap = "Heat map of number of reps per algal species collected by latitude.\\label{fig:algallat}"}
algal_heat <- filter(samples, kingdom %in% c('Plantae', 'Chromista'))
# for ice cover heat maps 
algal_heat_lat <- algal_heat %>%
  group_by(LAT_DD.1, genusSpecies) %>%
  summarize(sum(count))
# rename column
names(algal_heat_lat)[names(algal_heat_lat)=="sum(count)"] <- "totalSpecies"
ggplot(data = algal_heat_lat, aes(x = LAT_DD.1, y = genusSpecies)) +
  geom_tile(aes(fill = totalSpecies)) +
  scale_fill_viridis(option = "B", begin = 0.3, end = 1) +
  theme_dark() +
  theme(panel.background = element_rect(fill = 'black')) +
  theme(text = element_text(size = 22)) +
  geom_text(show.legend = FALSE, colour = "white", size = 8, aes(label = totalSpecies)) + 
  theme(axis.text.x = element_text(angle = 270))

```

```{r, echo = FALSE, fig.fullwidth = TRUE, warning = FALSE, message = FALSE, fig.height = 20, fig.width = 15, fig.cap = "Heat map of number of reps per algal species collected by ice cover category.\\label{fig:algalice}"}
algal_heat <- filter(samples, kingdom %in% c('Plantae', 'Chromista'))
# for ice cover heat maps 
algal_heat_cover <- algal_heat %>%
  group_by(IceCoverCat, genusSpecies) %>%
  summarize(sum(count))
# rename column
names(algal_heat_cover)[names(algal_heat_cover)=="sum(count)"] <- "totalSpecies"
ggplot(data = algal_heat_cover, aes(x = IceCoverCat, y = genusSpecies)) +
  geom_tile(aes(fill = totalSpecies)) +
  scale_fill_viridis(option = "B", begin = 0.3, end = 1) +
  theme_dark() +
  theme(panel.background = element_rect(fill = 'black')) +
  theme(text = element_text(size = 22)) +
  geom_text(show.legend = FALSE, colour = "white", size = 8, aes(label = totalSpecies)) + 
  theme(axis.text.x = element_text(angle = 270))

```

